# Stars

```{r}
library(sf)
library(stars)
library(tidyverse)
library(here)
```


## stars structure

- named list of arrays, each having a `dimension` attribute
- an attribute called `dimensions` of class `dimensions` that holds the dimensions metadata
- a class name that includes `stars`

## read data

> Monthly mean air temperature data from Geosphere

- https://dataset.api.hub.geosphere.at/app/frontend/raster/historical/spartacus-v2-1m-1km

```{r}
#| warning: false
#| message: false

# spartacus temp data
au_temp_jan_mar = read_ncdf(here("data/geodata/SPARTACUS - Spatial Dataset for Climate in Austria Datensatz_202301_202303.nc")) %>% st_set_crs("EPSG:3416")

# vienna for aggregation
vienna = read_sf(here("data/geodata/gems_oe_31287.fgb")) %>% dplyr::filter(str_detect(g_id, "^9.*")) %>% st_transform(3416)
plot(au_temp_jan_mar)
```
## Find mean temperature for Jan, Feb and March for each District in Vienna

```{r}
per_district = aggregate(au_temp_jan_mar, vienna, FUN=mean)
plot(per_district)
```

## In Which month was the max temp for each district

```{r}
st_apply(au_temp_jan_mar, vienna, which.max)
```

## For each value in the timedimension find the max cell (with geocoordinates)

```{r}
# start with one
jan = au_temp_jan_mar[,,,1]
# find the max cell and its centroid
which.max(jan[[1]])


```



## Dimensions

- Dimensions is a named list

```{r}
dimenstions_attr = attr(au_temp_jan_mar, "dimensions")
dimenstions_attr
```

## Attributes




